% with font size 10, I think there is just too much Information per page.
\documentclass[a4paper, 12pt]{scrreprt}

% border sizes
\setlength{\topmargin}{0.0in}
\setlength{\oddsidemargin}{0.33in}
\setlength{\textheight}{9.0in}
\setlength{\textwidth}{6.0in}

%fonts
% times is the font I (Mirco) thinks looks the most calm, ordered and math-textbook like
\usepackage{times}
% use times style for mathmode, too
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% integer long division
\usepackage{longdivision}
% polynomial long division
\usepackage{polynom}

\usepackage{amssymb,amsmath,amsthm,mathrsfs,xspace,multicol}
\usepackage{polydiv}
\usepackage{amscd}
\usepackage[all]{xy}

\usepackage{algorithm}
\usepackage{algpseudocode}
% We use algorithmicX, since we need to customize commands (\algblockdefx)
% See https://ftp.agdsn.de/pub/mirrors/latex/dante/macros/latex/contrib/algorithmicx/algorithmicx.pdf for howTo customize
\usepackage{algorithmicx}


\usepackage{enumitem} % for personalizing enumeration with itemize

% the sagetex evironment
\usepackage{sagetex}
\usepackage{xcolor}
% color of the sagecommandline-tex-environment
\lstset{language=Sage,
commentstyle={\ttfamily\color{black}},
keywordstyle={\ttfamily\color{black}\bfseries},
stringstyle ={\ttfamily\color{black}\bfseries},
tabsize = 4,
basicstyle={\small \ttfamily},
backgroundcolor= \color{white},
}


%Load the following packages last
\usepackage{hyperref} % for all kinds of linke
\usepackage{cleveref}%must be loaded after `hyperref`


%margins
\usepackage[a4paper, margin=2.5cm]{geometry}


%bibliography management
\usepackage{natbib}
\bibliographystyle{unsrtnat}%we can discuss what style we want to use

%math typesetting
\usepackage{amsfonts}

% Table of Contents

\setcounter{tocdepth}{4}

%	Macros
% Fields and categories
\newcommand{\C}{\mathbb{C}}
\newcommand{\G}{\mathbb{G}}
\newcommand{\Prim}{\mathbb{P}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\NN}{\mathbb{N}_0}
\newcommand{\K}{\mathbb{K}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\X}{\mathfrak{X}}
\newcommand{\Oinf}{\mathcal{O}}

\newcommand{\Zdiv}[2]{#1 \text{ div } #2}
\newcommand{\Zmod}[2]{#1 \text{ mod } #2}
\newcommand{\kongru}[3]{#1 \equiv #2 \quad \text{( mod } #3 \text{ )}}

% inline code
\newcommand{\icode}[1]{\lstinline{#1}}

% theorem styles
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[subsection]
\newtheorem{prop}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{conjecture}{Conjecture}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation and Symbols}

% Examples get their own counter
\newcounter{ExampleCounter}
\newtheorem{example}[ExampleCounter]{Example}

% Exercises get their own counter
\newcounter{ExerciseCounter}
\newtheorem{exercise}[ExerciseCounter]{Exercise}

% Jargon get their own counter
\newcounter{JargonCounter}
\newtheorem{jargon}[JargonCounter]{Jargon}

% algorithmicx commands
%\algnewcommand\algorithmicinput{\textbf{Statement:}}
%\algnewcommand\Statement{\item[\algorithmicinput]}

\title{Moonmath manual}
\author{TechnoBob and the Least Scruples crew}
\date{\today}

\begin{document}

\maketitle

Lorem \textbf{ipsum} dolor sit amet, consectetur adipiscing elit. Pellentesque semper viverra dictum.  Fusce interdum venenatis leo varius vehicula. Etiam ac massa dolor. Quisque vel massa faucibus, facilisis nulla nec, egestas lectus. Sed orci dui, egestas non felis vel, fringilla pretium odio. \textit{Aliquam} vel consectetur felis. Suspendisse justo massa, maximus eget nisi a, maximus gravida mi.

Here is a citation for demonstration: \cite{lamport1982the}

\include{chapters/intro-moonmath}
\include{chapters/zoo-moonmath}
\include{chapters/preliminaries-moonmath}
\include{chapters/arithmetics-moonmath}
\include{chapters/algebra-moonmath}
\include{chapters/elliptic-curves-moonmath}
\include{chapters/computational-models-moonmath}
\include{chapters/exercise-solutions-moonmath}

\bibliography{moonmath}%use `moonmath.bib` to look for references, and print bibliography at this point in the document

\end{document}
