\chapter*{Preface}\addcontentsline{toc}{chapter}{Preface}

This book began as a set of lectures and notes the author gave at the \href{https://www.zksummit.com/}{Zero Knowledge Summit – ZK0x02} in Berlin. It arose from the desire to collect the scattered information around the topic of zk-SNARKS and present them to an audience that does not have a strong background in cryptography.

After more than a decade of intense and fast-paced research on zk-SNARKs by mathematicians
and cryptographers around the globe, the field is now racing towards full maturity, and we believe we can see the first saturation effects appear at the horizon.\sme{what does saturation mean here?} We hope this book will equip the reader with most
of the grounding knowledge they need to tackle the vast literature in this
remarkable field.

As security auditors at Least Authority, we have audited numerous SNARK-based systems. This book includes a many of our learnings from the time we spent on various audits, and serves as a polished version of what one might call an auditor's adaptation of classical \href{https://en.wikipedia.org/wiki/Lab_notebook}{lab notebooks}.  

We intend to let illustrative examples drive the discussion and present the key ideas of all basic concepts relevant to the understanding of zk-SNARKS with as little mathematics as possible. 

For those who are new to this topic, it is our hope that this book might be particularly useful as a first read and prelude to more complete or advanced expositions.

For more advanced
readers and those who want to deepen their knowledge, we have provided reading recommendations throughout the manual.


